19Q2

### 2019-07-08 manuscript_old.Rmd

---
output              :
  pdf_document      :
    toc             : true
    toc_depth       : 2
    number_sections : true
    fig_caption     : true
    df_print        : kable
    highlight       : tango
    # citation_package: natbib
title               : "Manuscript Draft"
author              :
- Sally E. Claridge
- Daniel M. Charytonowicz
- Adam A. Margolin
date                : "`r format(Sys.time(), '%d %B %Y')`"
bibliography        : manuscript.bib
fontsize            : 12pt
geometry            : margin=1in
linkcolor           : blue
urlcolor            : blue
citecolor           : blue
abstract            : Multiple high-throughput functional screens in cancer cell lines have generated large amounts of information on drug efficacy in a variety of genomic contexts and cancer lineages. Inconsistencies between datasets have led us to investigate whether these large-scale screens reproduce established clinical drug-gene associations and if genomic features particular to specific genes improve said reproducibility. We evaluated three large-scale, small-molecule drug screens within the context of clinical interpretations derived from a new cancer variant annotation resource published by the Variant Interpretation for Cancer Consortium (VICC). We identified low levels of concordance between the three drug screen datasets and clinical drug-gene associations using mutation status, gene expression, and copy number as genomic indicators. Less than half of the clinical drug-gene cancer associations from the VICC resource were identified in these three drug screen datasets, suggesting a barrier to translating findings from these large-scale screens into the clinic.
---

```{r setup, include=FALSE}
library(ggpubr)
library(tidyverse)
library(kableExtra)

# install.packages("tinytex")
# tinytex::install_tinytex()
# install.packages("citr")
# devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
# update wordcount in Addins
```

```{r analysis-preferences, include=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

```{r load-processed-data, echo=FALSE, include=FALSE}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info_edit_20190708.csv", sep = ",", header = TRUE, row.names = 1)
maf_df <- readRDS("./data_munging/19Q2/maf_df_19Q2.rds")
maf_changes_df <- readRDS("./data_munging/19Q2/maf_changes_df_19Q2.rds")
cn_melt <- readRDS("./data_munging/19Q2/cn_melt_19Q2.rds")
ge_melt <- readRDS("./data_munging/19Q2/ge_melt_19Q2.rds")
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions_edit_20190708.csv.gz", header = TRUE, sep = ",")

ccle <- readRDS("./data_munging/19Q2/ccle_data_19Q2.rds") %>% drop_na(Drug)
ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q2.rds") %>% drop_na(Drug)
gdsc <- readRDS("./data_munging/19Q2/gdsc_data_19Q2.rds") %>% drop_na(Drug)
fus <- readRDS("./data_munging/rds/fusions_merged.rds") %>% drop_na(Drug)
fus_summary <- fus %>% count(Database, Drug, FusionPair, Fusion_Status)
prot <- readRDS("./data_munging/rds/protein_changes_merged.rds")

crispr <- readRDS("./data_munging/19Q2/crispr_data_19Q2.rds")

crispr_prot <- readRDS("./data_munging/rds/protein_changes_merged.rds")
```

@@

After filtering, `r nrow(g2p_filt)` level-A G2P interpretations were carried forward in subsequent analyses, constituting `r length(g2p_drugs)` unique drugs and `r length(g2p_genes)` unique genes in `r length(g2p_druggenes)` distinct level-A combinations (*Figure 1A: Pink heatmap*). The number of level-A interpretations per drug ranged from `r range(g2p_drug_tally$n)[1]` to `r range(g2p_drug_tally$n)[2]` ($M =$ `r formatC(mean(g2p_drug_tally$n), digits = 2)`, $SD =$ `r formatC(sd(g2p_drug_tally$n), digits = 2)`), `r range(g2p_gene_tally$n)[1]` to `r range(g2p_gene_tally$n)[2]` ($M =$ `r formatC(mean(g2p_gene_tally$n), digits = 2)`, $SD =$ `r formatC(sd(g2p_gene_tally$n), digits = 2)`) per gene, and for each level-A drug-gene combination, the range was `r range(g2p_druggene_tally$n)[1]` to `r range(g2p_druggene_tally$n)[2]` ($M =$ `r formatC(mean(g2p_druggene_tally$n), digits = 2)`, $SD =$ `r formatC(sd(g2p_druggene_tally$n), digits = 2)`). These unique drugs, genes, and combinations are hereafter referred to as G2P drugs, G2P genes, and G2P associations. Filtering our screening datasets by these clinical criteria yields panels of potentially clinically relevant omic features. Of the `r length(g2p_drugs)` unique G2P drugs, N of them were screened in at least one of the drug screen datasets, N drugs were screened in at least two of the three datasets, and only N were screened in all three datasets. The remaining N un-screened G2P drugs suggests an incompleteness of these drug screens, though these projects frequently release new screening data that may include these drugs in future releases.

> Mutation status

> Fusions

> Copy number

> Gene expression

@@

@@

@@

## Drug screen dataset retrieval

**Need to update with Poz info**

@@

@@

>>>>

---
title             : "Manuscript Draft"
shorttitle        : "Subtitle"

author: 
  - name          : "Sally E. Claridge"
    affiliation   : "1"
  - name          : "Daniel M. Charytonowicz"
    affiliation   : "1"
  - name          : "Adam A. Margolin"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Icahn School of Medicine at Mount Sinai, Deparment of Genetics and Genomic Sciences, One Gustave L. Levy Place, Box 1498, New York, NY 10029"
    email         : "adam.margolin@mssm.edu"

affiliation:
  - id            : "1"
    institution   : "Department of Genetics and Genomic Sciences, Icahn Institute for Data Science and Genomic Technology, Icahn School of Medicine at Mount Sinai, New York, NY"
  - id            : "2"
    institution   : "Cancer Target Discovery and Development Network (Oregon Health and Science University, Portland, OR), National Cancer Institute, National Institutes of Health"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Multiple high-throughput functional screens in cancer cell lines have generated large amounts of information on drug efficacy in a variety of genomic contexts and cancer lineages. Inconsistencies between datasets have led us to investigate whether these large-scale screens reproduce established clinical drug-gene associations and if genomic features particular to specific genes improve said reproducibility. We evaluated three large-scale, small-molecule drug screens and one CRISPR/Cas9 gene essentiality screen, all within the context of clinical interpretations derived from a new cancer variant annotation resource published by the Variant Interpretation for Cancer Consortium (VICC). We identified low levels of concordance between the three drug screen datasets and clinical drug-gene associations using mutation status, gene expression, and copy number as genomic indicators. Less than half of the clinical drug-gene cancer associations from the VICC resource were identified in these three drug screen datasets, suggesting a barrier to translating findings from these large-scale screens into the clinic.

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["manuscript.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
biblio-style      : "apa"
output            : papaja::apa6_pdf
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(papaja)
library(bookdown)
library(VennDiagram)
library(lettercase)
# install.packages("tinytex")
# tinytex::install_tinytex()
# install.packages("citr")
# devtools::install_github("benmarwick/wordcountaddin", type = "source", dependencies = TRUE)
# update wordcount in Addins
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

```{r drug-data, echo=FALSE, include=FALSE}
ccle_data_ptmuts <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes_ptmuts.rds")
ctrp_data_ptmuts <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes_ptmuts.rds")
gdsc_data_ptmuts <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes_ptmuts.rds")

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ccle_data_g2p$Mutation_Status_Nonsilent <- factor(ccle_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ctrp_data_g2p$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_druggenes)
gdsc_data_g2p$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drug-gene-lineage associations
ccle_data_g2p_dgl <- filter(ccle_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ccle_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_dgl <- filter(ctrp_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ctrp_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_dgl <- filter(gdsc_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
gdsc_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drugs and genes
ccle_data_g2p_grid <- filter(ccle_data_ptmuts, Drug %in% g2p_drugs)
ccle_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_grid <- filter(ctrp_data_ptmuts, Drug %in% g2p_drugs)
ctrp_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_grid <- filter(gdsc_data_ptmuts, Drug %in% g2p_drugs)
gdsc_data_g2p_grid$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Wilcoxon test results
ccle_signif_g2p <- readRDS("./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_g2p_exp <- e2exp_converter(ccle_signif_g2p$FDR_cutoff[1], p_type = "pk")
ccle_signif_g2p_grid <- readRDS("./data_munging/rds/ccle_signif_g2p_grid.rds")
ccle_signif_g2p_grid_exp <- e2exp_converter(ccle_signif_g2p_grid$FDR_cutoff[1], p_type = "pk")
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
ccle_signif_g2p_lineage_exp <- e2exp_converter(ccle_signif_g2p_lineage$FDR_cutoff[1], p_type = "pk")
ccle_signif_grid_lineage <- readRDS("./data_munging/rds/ccle_signif_grid_lineage.rds")
ccle_signif_grid_lineage_exp <- e2exp_converter(ccle_signif_grid_lineage$FDR_cutoff[1], p_type = "pk")

ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")
ctrp_signif_g2p_exp <- e2exp_converter(ctrp_signif_g2p$FDR_cutoff[1], p_type = "pk")
ctrp_signif_g2p_grid <- readRDS("./data_munging/rds/ctrp_signif_g2p_grid.rds")
ctrp_signif_g2p_grid_exp <- e2exp_converter(ctrp_signif_g2p_grid$FDR_cutoff[1], p_type = "pk")
ctrp_signif_g2p_lineage <- readRDS("./data_munging/rds/ctrp_signif_g2p_lineage.rds")
ctrp_signif_g2p_lineage_exp <- e2exp_converter(ctrp_signif_g2p_lineage$FDR_cutoff[1], p_type = "pk")
ctrp_signif_grid_lineage <- readRDS("./data_munging/rds/ctrp_signif_grid_lineage.rds")
ctrp_signif_grid_lineage_exp <- e2exp_converter(ctrp_signif_grid_lineage$FDR_cutoff[1], p_type = "pk")

gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")
gdsc_signif_g2p_exp <- e2exp_converter(gdsc_signif_g2p$FDR_cutoff[1], p_type = "pk")
gdsc_signif_g2p_grid <- readRDS("./data_munging/rds/gdsc_signif_g2p_grid.rds")
gdsc_signif_g2p_grid_exp <- e2exp_converter(gdsc_signif_g2p_grid$FDR_cutoff[1], p_type = "pk")
gdsc_signif_g2p_lineage <- readRDS("./data_munging/rds/gdsc_signif_g2p_lineage.rds")
gdsc_signif_g2p_lineage_exp <- e2exp_converter(gdsc_signif_g2p_lineage$FDR_cutoff[1], p_type = "pk")
gdsc_signif_grid_lineage <- readRDS("./data_munging/rds/gdsc_signif_grid_lineage.rds")
gdsc_signif_grid_lineage_exp <- e2exp_converter(gdsc_signif_grid_lineage$FDR_cutoff[1], p_type = "pk")

# Mut summaries
drug_grid_shared_dg <- union(ccle_signif_g2p_grid[!is.na(ccle_signif_g2p_grid$p.signif.adj),]$Drug_Gene, union(ctrp_signif_g2p_grid[!is.na(ctrp_signif_g2p_grid$p.signif.adj),]$Drug_Gene, gdsc_signif_g2p_grid[!is.na(gdsc_signif_g2p_grid$p.signif.adj),]$Drug_Gene))

# CCLE: GE/CN correlations
ccle_cor_res <- ccle_data_g2p_grid %>% group_by(Drug_Gene) %>% summarize(
  Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$estimate,
  Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$p.value,
  Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$estimate,
  Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$p.value,
  Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate,
  Spearman_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value,
  Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate,
  Spearman_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value)
ccle_cor_res <- merge(g2p_druggene_tally, ccle_cor_res, by = "Drug_Gene", all.y = TRUE)
ccle_cor_res$n <- ifelse(is.na(ccle_cor_res$n), 0, ccle_cor_res$n)
ccle_cor_res$InG2P <- ifelse(ccle_cor_res$n != 0, "Yes", "No")
ccle_cor_res$Dataset <- "CCLE"

# CTRP: GE/CN correlations
ctrp_cor_res <- ctrp_data_g2p_grid %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value)
ctrp_cor_res <- merge(g2p_druggene_tally, ctrp_cor_res, by = "Drug_Gene", all.y = TRUE)
ctrp_cor_res$n <- ifelse(is.na(ctrp_cor_res$n), 0, ctrp_cor_res$n)
ctrp_cor_res$InG2P <- ifelse(ctrp_cor_res$n != 0, "Yes", "No")
ctrp_cor_res$Dataset <- "CTRP"

# GDSC: GE/CN correlations
gdsc_cor_res <- gdsc_data_g2p_grid %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "spearman", use = "complete.obs", exact = FALSE)$p.value, Spearman_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$estimate, Spearman_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "spearman", use = "complete.obs", exact = FALSE)$p.value)
gdsc_cor_res <- merge(g2p_druggene_tally, gdsc_cor_res, by = "Drug_Gene", all.y = TRUE)
gdsc_cor_res$n <- ifelse(is.na(gdsc_cor_res$n), 0, gdsc_cor_res$n)
gdsc_cor_res$InG2P <- ifelse(gdsc_cor_res$n != 0, "Yes", "No")
gdsc_cor_res$Dataset <- "GDSC"

# Dataset counts
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)
for(drug in g2p_drugs) {
  counter = 0
  if(drug %in% as.character(unique(ccle_data_g2p$Drug))) { counter <- counter + 1 }
  if(drug %in% as.character(unique(ctrp_data_g2p$Drug))) { counter <- counter + 1 }
  if(drug %in% as.character(unique(gdsc_data_g2p$Drug))) { counter <- counter + 1 }
  drug_summ <- rbind(drug_summ, data.frame(Drug = drug, Dataset_Count = counter))
}
```

```{r table-g2p-nonCID, results="asis"}
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize("CID" = DrugID, "N_G2P" = n, "%" = printnum(n / nrow(g2p_nonCID) * 100))
colnames(summ)[colnames(summ) == "Drug"] <- "Compound"
summ$Compound[summ$Compound == "egfr"] <- "EGFR"
summ$Compound[summ$Compound == "mab"] <- "mAb"
summ <- summ[order(summ$N_G2P),]

apa_table(summ, align = "c", font_size = "small", caption = "Level-A G2P interpretations excluded from analyses", note = "N_G2P (%) = number of associated G2P interpretations, % is out of column total; mAb = monoclonal antibody; NA indicates G2P interpretation does not have an associated compound.")
```

@@

After filtering, `r nrow(g2p_filt)` level-A G2P interpretations were carried forward in subsequent analyses, constituting `r length(g2p_drugs)` unique drugs and `r length(g2p_genes)` unique genes in `r length(g2p_druggenes)` distinct level-A combinations (*Figure 1A: Pink heatmap*). The number of level-A interpretations per drug ranged from `r range(g2p_drug_tally$n)[1]` to `r range(g2p_drug_tally$n)[2]` ($M =$ `r printnum(mean(g2p_drug_tally$n))`, $SD =$ `r printnum(sd(g2p_drug_tally$n))`), `r range(g2p_gene_tally$n)[1]` to `r range(g2p_gene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_gene_tally$n))`, $SD =$ `r printnum(sd(g2p_gene_tally$n))`) per gene, and for each level-A drug-gene combination, the range was `r range(g2p_druggene_tally$n)[1]` to `r range(g2p_druggene_tally$n)[2]` ($M =$ `r printnum(mean(g2p_druggene_tally$n))`, $SD =$ `r printnum(sd(g2p_druggene_tally$n))`). These unique drugs, genes, and combinations are hereafter referred to as G2P drugs, G2P genes, and G2P associations. Filtering our screening datasets by these clinical criteria yields panels of potentially clinically relevant omic features. Of the `r length(g2p_drugs)` unique G2P drugs, `r nrow(drug_summ[drug_summ$Dataset_Count >= 1,])` of them were screened in at least one of the drug screen datasets, `r nrow(drug_summ[drug_summ$Dataset_Count >= 2,])` drugs were screened in at least two of the three datasets, and only `r nrow(drug_summ[drug_summ$Dataset_Count >= 3,])` were screened in all three datasets. The remaining `r nrow(drug_summ[drug_summ$Dataset_Count == 0,])` un-screened G2P drugs suggests an incompleteness of these drug screens, though these projects frequently release new screening data that may include these drugs in future releases.

Using cell-line-specific point mutation annotations from the CCLE, we compared z-score-transformed area under the dose-response curve (AUC) distributions between mutant and wildtype cell lines for all G2P associations tested in the CCLE, CTRP, and GDSC drug screens. Of the `r length(g2p_druggenes)` G2P associations, data was available for `r length(ccle_signif_g2p$Drug_Gene)` associations in CCLE, `r length(ctrp_signif_g2p$Drug_Gene)` in CTRP, and `r length(gdsc_signif_g2p$Drug_Gene)` in GDSC. Controlling for a false discovery rate (FDR) < 5%, CCLE yielded no significant correlations (Wilcoxon rank-sum tests, `r ccle_signif_g2p_exp`), though the most significant G2P association was `r drugstring2intext_converter(ccle_signif_g2p$Drug_Gene[1], "dg")` (`r e2exp_converter(ccle_signif_g2p$p.adj[1], p_type = "padj")`). In CTRP, AUC from `r nrow(filter(ctrp_signif_g2p[!is.na(ctrp_signif_g2p$p.signif.adj),]))` drug-gene associations correlated with the gene's mutation status (`r ctrp_signif_g2p_exp`): `r drugstring2intext_converter(ctrp_signif_g2p$Drug_Gene[1], "dg")` (`r e2exp_converter(ctrp_signif_g2p$p.adj[1], p_type = "padj")`), `r drugstring2intext_converter(ctrp_signif_g2p$Drug_Gene[2], "dg")` (`r e2exp_converter(ctrp_signif_g2p$p.adj[2], p_type = "padj")`), and `r drugstring2intext_converter(ctrp_signif_g2p$Drug_Gene[3], "dg")` (`r e2exp_converter(ctrp_signif_g2p$p.adj[3], p_type = "padj")`). `r str_ucfirst(drugstring2intext_converter(gdsc_signif_g2p$Drug_Gene[1], "dg"))` (`r e2exp_converter(gdsc_signif_g2p$p.adj[1], p_type = "padj")`) was the only significant association in GDSC (`r gdsc_signif_g2p_exp`).

> **(???) Describe why we compare all pairwise combos rather than filtering for just the drug-gene combos in G2P?**

To test for potentially novel drug-gene associations, we conducted similar Wilcoxon rank-sum tests for all pairwise combinations of G2P genes and drugs tested in the three drug screens (*Figures 1B: Green heatmaps*). At an unadjusted $\alpha = 0.05$, the tests revealed that in CCLE, AUC significantly correlated with mutation status for `r nrow(filter(ccle_signif_g2p_grid, p < 0.05))` out of `r nrow(ccle_signif_g2p_grid)` (`r printnum(nrow(filter(ccle_signif_g2p_grid, p < 0.05)) / nrow(ccle_signif_g2p_grid) * 100)`%) tested drug-gene combinations were significantly different, `r nrow(filter(ctrp_signif_g2p_grid, p < 0.05))` out of `r nrow(ctrp_signif_g2p_grid)` (`r printnum(nrow(filter(ctrp_signif_g2p_grid, p < 0.05)) / nrow(ctrp_signif_g2p_grid) * 100)`%) were significantly different for CTRP, and `r nrow(filter(gdsc_signif_g2p_grid, p < 0.05))` out of `r nrow(gdsc_signif_g2p_grid)` (`r printnum(nrow(filter(gdsc_signif_g2p_grid, p < 0.05)) / nrow(gdsc_signif_g2p_grid) * 100)`%) for GDSC.

Controlling for a FDR < 5%, `r drugstring2intext_converter(ccle_signif_g2p_grid$Drug_Gene[1], "dg")` (`r e2exp_converter(ccle_signif_g2p_grid$p.adj[1], p_type = "padj")`) was the only significant drug-gene combination tested in CCLE (`r ccle_signif_g2p_grid_exp`). CTRP (`r ctrp_signif_g2p_grid_exp`) yielded `r nrow(ctrp_signif_g2p_grid[!is.na(ctrp_signif_g2p_grid$p.signif.adj),])` significant combinations, one of which was selumetinib-*BRAF*. There were `r nrow(gdsc_signif_g2p_grid[!is.na(gdsc_signif_g2p_grid$p.signif.adj),])` significant G2P associations in the GDSC dataset (`r gdsc_signif_g2p_grid_exp`). Of all the significant correlations between mutation status and AUC across datasets, the only genes represented were *KRAS* and *BRAF* (Figure\ \@ref(fig:drug-venn)). Of the `r length(drug_grid_shared_dg)` significant drug-gene associations across datasets, `r length(intersect(drug_grid_shared_dg, g2p_druggenes))` were G2P associations. The remaining `r length(drug_grid_shared_dg) - length(intersect(drug_grid_shared_dg, g2p_druggenes))` associations were `r drugstring2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[1], "dg")` and `r drugstring2intext_converter(setdiff(drug_grid_shared_dg, g2p_druggenes)[2], "dg")`.

> **(???) Trametinib and selumetinib are both *MEK* inhibitors, and pazopanib is a *VEGFR* inhibitor. VEGF signaling can influence the MAPK/ERK signaling pathway, which promotes cell proliferation and survival. This pathway includes *RAS*, *BRAF*, and *MEK* proteins.**

@@

(ref:drug-venn) **Overlap of significant drug-gene associations from CCLE, CTRP, and GDSC**. Overlap of drug-gene associations drawn from all pairwise combinations of G2P drugs and G2P genes whose drug effects significantly correlated with (**A**) mutation status, (**B**) gene expression, and (**C**) copy number across CCLE, CTRP, and GDSC.

```{r drug-venn, echo = FALSE, message=FALSE, warning=FALSE, fig.cap='(ref:drug-venn)'}
knitr::include_graphics("./plots/manuscript/drugscreen_venns.pdf")
```

> **(???) Cut these next three paragraphs since we cut the analysis, but keep for future reference**

We compared the distribution of level-A G2P associations (*Figure 1A*) and the associated p-values from the mutation status comparisons (*Figure 1B*) with a null distribution of p-values derived from all pairwise combinations of G2P genes and the G2P drugs tested in each dataset, less any of the 88 distinct combinations with G2P interpretations, i.e. the non-gray squares in the right heatmap of *Figure 1A (Figure 3A*). To formally test concordance between clinically actionable variants and drug screen results, we conducted an independent, two-group t-test to compare the distributions of p-values of drug-gene combinations that either were or were not represented with level-A evidence in G2P, which indicated that the distributions of p-values did not significantly differ for CCLE [t(9) = 0.25, p = 0.81], CTRP [t(51) = 0.85, p = 0.40], nor GDSC [t(47) = -0.56, p = 0.58]  (Table 2). This further supports that mutation status is not a robust genomic feature for predicting therapeutic responses.

Similar tests were conducted for the distributions of Pearson correlation coefficients (r) between the number of level-A G2P associations and gene expression (RPKM) and copy number (log2 ratio) of the gene in the association (*Figures 3B-C*). Example r derivations are shown in *Figures 2B-C*. Absolute values of r were used to simplify extreme values in both directions for testing. The difference in the CCLE |r| distributions for both gene expression [t(8) = 1.43, p = 0.19] and copy number [t(8) = 2.48, p = 0.23] were non-significant, which is likely explained by the small sample size since only five G2P drugs were screened in the CCLE dataset. For both the CTRP and GDSC screens, |r| values for gene expression correlations were significantly greater for the level-A G2P associations than the null distribution [tCTRP(49) = 3.79, pCTRP < 0.001; tGDSC(44) = 2.97, pGDSC = 0.005]. The |r| distributions using copy number were also significantly different for CTRP [t(49) = 2.48, p = 0.017] but not for GDSC [t(44) = 1.95, p = 0.058].

In all three datasets for both level-A G2P associations and null distribution associations, there are more significant |r| values when using gene expression as the genomic feature than when using copy number (Table 2), which corresponds with previous work showing that gene expression is an informative molecular feature when assessing drug sensitivity [@jang2014; @sirota2011]. Additionally, in all cases, there are more significant r values when using copy number than there are significant p-values from the mutation status comparisons. Though observational, these results suggest an order of predictive robustness for these three genomic features when used in isolation: Gene expression > copy number > mutation status. Additionally, across all three genomic features, CTRP consistently identified a higher percentage of its possible level-A G2P associations than GDSC, and similarly, GDSC identified a higher percentage than CCLE. In the CCLE-GDSC debate, inconsistency between the datasets was attributed to the groups using two very different assays for generating AUC measurements though it was undecided which was more accurate [@haibe-kains2013; @jang2014; @weinstein2013]. Our results provide very preliminary evidence that GDSC is more accurate than CCLE in regards to reproducing clinical associations. However, CTRP may be even more accurate than both GDSC and CCLE.

@@

In the level-A G2P interpretations, there were `r nrow(g2p_dgl_tally)` unique drug-gene-lineage combinations (DGLs), representing seven of our curated cancer lineages: `r paste(g2p_lins[1:length(g2p_lins) -1], collapse = ", ")`, and `r g2p_lins[length(g2p_lins)]`. CCLE had data for `r length(unique(ccle_data_g2p_dgl$Drug_Gene_Lin))` DGLs, CTRP had data for `r length(unique(ctrp_data_g2p_dgl$Drug_Gene_Lin))`, and GDSC had data for `r length(unique(gdsc_data_g2p_dgl$Drug_Gene_Lin))`, yielding a total of `r length(union(union(unique(ccle_data_g2p_dgl$Drug_Gene_Lin), unique(ctrp_data_g2p_dgl$Drug_Gene_Lin)), unique(gdsc_data_g2p_dgl$Drug_Gene_Lin)))` unique DGLs across all three datasets. Only `r length(intersect(intersect(unique(ccle_data_g2p_dgl$Drug_Gene_Lin), unique(ctrp_data_g2p_dgl$Drug_Gene_Lin)), unique(gdsc_data_g2p_dgl$Drug_Gene_Lin)))` DGLs were screened in all three datasets.

Prior to correction for an FDR of 5%, 1 DGL, `r drugstring2intext_converter(ccle_signif_g2p_lineage$Drug_Gene_Lin[1], "dgl")`, was identified in the CCLE screen when using mutation status as a correlate ($n =$ `r nrow(ccle_signif_g2p_lineage)`, `r length(unique(ccle_data_g2p_dgl$Drug_Gene_Lin)) - nrow(ccle_signif_g2p_lineage)` DGLs were not analyzed due to lack of mutation annotations in *UGT1A1* in "cancer" lineages), but CTRP ($n =$ `r nrow(ctrp_signif_g2p_lineage)`) and GDSC ($n =$ `r nrow(gdsc_signif_g2p_lineage)`) identified `r nrow(filter(ctrp_signif_g2p_lineage, !is.na(p.signif)))` and `r nrow(filter(gdsc_signif_g2p_lineage, !is.na(p.signif)))`, respectively, in skin cancer, lung cancer, and cancer lineages. After Benjamini-Hochberg FDR correction, none of the datasets yielded significant DGL interactions.

> **(???) KS test results: Need to flesh out this section**

(Figure\ \@ref(fig:ksplot))

(ref:ksplot) **Comparison of drug-gene AUC vs. gene expression and vs. copy number in G2P and non-G2P cell lines**. Comparison of Spearman correlation coefficients ($r_s$) derived from copy number (log$_2$ ratio) vs. AUC z-scores and gene expression (RPKM) vs. AUC z-scores for G2P drug-gene associations. Correlation coefficients were grouped by whether or not the associated lineage for each comparison was in a level-A G2P drug-gene-lineage association or not. The distributions of these grouped coefficients were compared using Kolmogorov-Smirnov tests to identify whether correlations in G2P lineages were distinct from correlations in non-G2P lineages.

```{r ksplot, echo = FALSE, warning = FALSE, message = FALSE, fig.cap='(ref:ksplot)'}
knitr::include_graphics("./plots/manuscript/g2p_spearman_ks_res_plot.pdf")
```

@@

## Drug screen dataset retrieval and harmonization

@@

### AUC harmonization

The CCLE reports cell line sensitivity to small molecules as an activity area measured above the dose-response curve (ActArea), while both CTRP and GDSC report small molecule activity as area under the dose-response curve (AUC). Since ActArea is reported with a scale of 0 to 8, we converted CCLE’s ActArea values to AUC via

$$AUC=1- \frac{ActArea}{8}$$

All AUC values across the three drug screen datasets were z-score transformed:

$$z = \frac{AUC - \mu_{AUC}}{\sigma_{AUC}}$$

where AUC is the mean AUC across all drug-cell line combinations in the dataset and AUC is the standard deviation.

@@

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

Finally, the APA guidelines require a note at the start of the reference section that explains what an asterisk means. This note can be added at the end of the document as follows.

<div id = "refs"></div>
\endgroup
